\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-11-06 doc/flexiblesusy-paper.tex
Move to appendix sections on C++ code adaptation

%macros
\newcommand{\sarah}{SARAH\@\xspace}
\newcommand{\fs}{FlexibleSUSY\@\xspace}
\newcommand{\mathematica}{Mathematica\xspace}
\newcommand{\ESSM}{E$_6$SSM\@\xspace}
\newcommand{\code}[1]{\lstinline|#1|}  % inline source code
\newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}
\newcommand{\DRbar}{\textoverline{DR}\xspace}
\newcommand{\MSbar}{\textoverline{MS}\xspace}
\newcommand{\unit}[1]{\,\text{#1}}      % units
\newcommand{\userinput}{\text{input}}
\newcommand{\pole}{\text{pole}}
\newcommand{\Lagr}{\mathcal{L}}
\newcommand{\unity}{\mathbf{1}}
\newcommand{\figref}[1]{\figurename~\ref{#1}}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\tabref}[1]{\tablename~\ref{#1}}
\newcommand{\ptitle}[1]{\emph{#1}}
\renewcommand{\ptitle}[1]{}
\makeatletter
\lstnewenvironment{numlstlisting}[1][]{%
  \lstset{%
    #1,
    numbers=left,
    firstnumber=auto,
    numberstyle=\tiny\sffamily}%
  \csname\@lst @SetFirstNumber\endcsname
}{%
  \csname \@lst @SaveFirstNumber\endcsname
}
\makeatother
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\re}{Re}
\DeclareMathOperator{\im}{Im}
\def\at{\alpha_t}
\def\ab{\alpha_b}
\def\as{\alpha_s}
\def\atau{\alpha_{\tau}}
\def\oat{O(\at)}
\def\oab{O(\ab)}
\def\oatau{O(\atau)}
\def\oatab{O(\at\ab)}
\def\oatas{O(\at\as)}
\def\oabas{O(\ab\as)}
\def\oatababq{O(\at\ab + \ab^2)}
\def\oatqatababq{O(\at^2 + \at\ab + \ab^2)}
\def\oatasatq{O(\at\as + \at^2)}
\def\oatasabas{O(\at\as +\ab\as)}
\def\oatasabasatq{O(\at\as + \at^2 +\ab\as)}
\def\oatq{O(\at^2)}
\def\oabq{O(\ab^2)}
\def\oatauq{O(\atau^2)}
\def\oabatau{O(\ab \atau)}
\def\oatplusabsq{O((\at+\ab)^2)}
\def\oas{O(\as)}
\def\oatauqatab{O(\atau^2 +\ab \atau )}

@@

\subsection{Current limitations and future extensions}
Although we try to handle as many models as possible there are still
some limitations to what can be done. In it's current form \fs assumes
all couplings are are real, therefore it is limited to CP conserving
versions of SUSY models.   Although it is implicit in the
title we would like to stress that currently we cannot provide a
spectrum generator for non-SUSY models.  This also means that like
other spectrum generators it is difficult to reliably predict the mass
spectrum in extreme cases such as Split-SUSY
\cite{Wells:2003tf,ArkaniHamed:2004fb,Giudice:2004tc} where the mass
scales of the SUSY model are drastically split leading to very large
logarithms which are not resummed.  In such cases the tower of
effective theories offers the best possibility for a solution.
However for this to work first \fs must be extended to include
non-SUSY models.  Finally the gauge group of the model should be
semi-simple and contain the standard model gauge group as factor so
that we can match to the SM at low energies.

@@

\fs allows the user to add leading two-loop contributions to the
CP-even and CP-odd Higgs self-energies as well as to the CP-even Higgs
tadpoles.  For MSSM-like models (with two CP-even Higgs bosons, one
CP-odd Higgs boson, one neutral Goldstone boson) routines dor
caluclating these corrections will be generated by setting
\code{UseHiggs2LoopMSSM = True} in the model file and by defining the
effective $\mu$-term \code{EffectiveMu = \\[Mu]}.  This will add the
zero-momentum corrections of the order $O(y_t^4 + y_b^2 y_t^2 +
y_b^4)$, $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$, $O(y_\tau^4)$ from
\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}.
For NMSSM-like models (with three CP-even Higgs bosons, two CP-odd
Higgs bosons, one neutral Goldstone boson) the two-loop contributions
are generated by setting \code{UseHiggs2LoopNMSSM = True} and by
defining the effective $\mu$-term like \code{EffectiveMu = \\[Lambda]
  vS / Sqrt[2]}, for example.  This will add the zero-momentum
corrections of the order $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ from
\cite{Degrassi:2009yq} and also the MSSM contributions of the order
$O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$ as well
\cite{Brignole:2001jy,Dedes:2003km} which only represent a partial
correction for that order in the NMSSM, but can be a good
approximation when singlet mixing is very small\footnote{These
  corrections may be disabled in the SLHA file, as described in
  section \secref{sec:PoleMasses}.}.

@@

The CKM mixing matrix is currently set to unity and CP-violating

@@

(i) the calculated lightest CP-even Higgs pole mass, $m_h$, differs

@@

Since the Higgs mass is a very important measurement and the two-loop
corrections can be larger than the current experimental error
\cite{Degrassi:2009yq} we recommend to set these switches to
\code{True} in any MSSM-\footnote{In \fs a model is MSSM-like if (i)
  its superpotential is approximately given by the MSSM one, (ii) it
  implements $R$-parity conservation or something equivalent, and
  (iii) contains two CP-even and CP-odd Higgs bosons, where one CP-odd
  Higgs boson may be a Goldstone boson, all with an MSSM-like coupling
  to $t$, $b$ and $\tau$.} and NMSSM-like\footnote{In \fs a model is
  NMSSM-like if (i) its superpotential is approximately given by the
  NMSSM one, (ii) it implements $R$-parity conservation or something
  equivalent, and (iii) contains three CP-even and CP-odd Higgs
  bosons, where one or two CP-odd Higgs bosons may be Goldstone
  bosons, all with an NMSSM-like coupling to $t$, $b$ and $\tau$\@.
  Examples for NMSSM-like models are the USSM and the \ESSM.} model.
However in such models the user should still consider whether these
corrections are really the leading corrections in the model or if
there are other potentially large two-loop corrections which are
missing.  For models with a more extended Higgs sector we recommend
that the leading log two-loop corrections are estimated by
generalizing those of the MSSM or NMSSM.

@@

\subsubsection{Stacking models in a tower of effective theories}
\label{sec:tower construction}

@@

\subsubsection{Integrating custom-built C++ components}
\label{sec:integrating-custom-built}

@@

To check the correctness of \fs's generated spectrum generators
extensive unit testing against Softsusy's MSSM and NMSSM
implementations (both $Z_3$-invariant and $Z_3$-violating variants)
has been carried out.  These unit tests systematically compare all
tree-level mass matrices, EWSB equations, one- and two-loop
$\beta$-functions, one- and two-loop self-energies and one- and
two-loop tadpoles numerically for various MSSM and NMSSM parameter
points.  All tested expressions were found to agree within double
machine precision.\footnote{Due to the systematic and detailed tests
  several bugs in Softsusy, SARAH and \fs could be identified and
  corrected.}  Furthermore, the overall pole mass spectrum after the
full fixed-point iteration has finished has been compared, and was
found to agree at the sub-permille level.  The origin of the
sub-permille level difference between \fs and Softsusy is the
different determination of the weak mixing angle
$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ in the SUSY model in the
\DRbar\ scheme: \fs calculates
$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from $M_W$ and $M_Z$, as
described in \secref{sec:calculation-of-gauge-couplings}, while
Softsusy determines $\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from
the muon decay constant $G_\mu$.  The approach used in Softsusy
results in more precise \DRbar\ gauge couplings at the $M_Z$ scale,
because the muon decay constant is known with a higher accuracy than
the $W$-boson mass.  Furthermore, in Softsusy $3.5.0$ some three-loop
$\beta$-functions and two-loop threshold corrections can be enabled in
the MSSM to increase the accuracy of the RG running and the
determination of the \DRbar\ gauge and Yukawa couplings at $M_Z$
\cite{Allanach:2014nba}.  These corrections are not implemented in \fs
so far.  The complete set of unit tests is shipped with \fs and can be
found in the \code{test/} directory.  The tests can be run with the
command \code{make execute-tests}.  All unit tests are carried out
nightly in order to continuously check the correctness of the meta
code and the generated spectrum generators for the shipped models.
The nightly test results can be found at
\url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/test.xhtml}

One of \fs's design goals is a short run-time.  In this section we
demonstrate that this goal was achieved by comparing the run-time of
two different sets of CMSSM spectrum generators:
%
\begin{itemize}
\item \emph{Without sfermion flavour violation:} Disallowing sfermion
  flavour violation simplifies the calculation of the pole masses,
  because flavour-off-diagonal sfermion self-energy matrix elements
  don't need to be calculated.  Here we compare \fs's non-flavour
  violating CMSSM spectrum generator FlexibleSUSY-NoFV (version 1.0.0)
  against SPheno (version 3.2.4) and Softsusy (version 3.4.0).
%
\item \emph{With sfermion flavour violation:} Allowing for sfermion
  flavour violation in general increases the run-time of spectrum
  generators, because the full $6\times 6$ sfermion self-energy
  matrices have to be calculated.  Here we compare FlexibleSUSY-FV
  (version 1.0.0) and SPhenoMSSM (generated with SARAH 4.1.0 and
  linked against SPheno 3.2.4).  Both spectrum generators are based on
  SARAH's MSSM model file, which allows for sfermion flavour
  violation.
\end{itemize}
%
For the run-time comparison the CKM matrix is set to unity, all CP
violating phases are set to zero and R-parity violation is disabled.
\fs and Softsusy are compiled with g++ 4.8.0 and Intel ifort 13.1.3
20130607.  SPheno and SPhenoMSSM are compiled with Intel ifort 13.1.3
20130607.\footnote{Intel's ifort compiler decreases the run-time of
  SPheno and SPhenoMSSM by approximately a factor $1.5$, compared to
  gfortran.}  We're generating $2\cdot 10^{4}$ random CMSSM parameter
points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
[50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
and $A_0\in [-1000,$ $1000]\unit{GeV}$.  For each point an SLHA input
file is created by appending the values of $m_0$, $m_{1/2}$,
$\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
SLHA template file given in \ref{sec:speed-test-slha-template-file}.
The resulting SLHA input file is passed to each spectrum generator and
the (wall-clock) time is measured until the program has finished.  The
average run-times for three different CPU types can be found in
\tabref{tab:run-time-comparison}.  The first column shows the run-time
on a Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
enabled.  The second column shows the run-time on the same processor
where both cores were enabled.  In the third column a machine with two
Intel Xeon CPUs (L5640, $2.27\unit{GHz}$, $6$ cores) was used.
%
\begin{table}[tbh]
  \centering
  \begin{tabular}{llll}
    \toprule
                            & Intel Core2 Duo    & Intel Core2 Duo   & $2$ $\times$ Intel Xeon\\
                            & (P8600, $1$ core)  & (P8600, $2$ cores)& (L5640, $6$ cores)\\
    \midrule
    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
    \midrule
    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
    SPhenoMSSM 4.1.0        & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
    \bottomrule
  \end{tabular}
  \caption{Average run-time of CMSSM spectrum generators
    for random parameter points.  The first three rows show three
    spectrum generators with disabled sfermion flavour violation.  Rows
    $4$--$5$ show two spectrum generators with enabled sfermion flavour violation, both based
    on SARAH's MSSM model file.}
  \label{tab:run-time-comparison}
\end{table}

>>>>

%macros
\newcommand{\sarah}{SARAH\@\xspace}
\newcommand{\fs}{FlexibleSUSY\@\xspace}
\newcommand{\mathematica}{Mathematica\xspace}
\newcommand{\ESSM}{E$_6$SSM\@\xspace}
\newcommand{\code}[1]{\lstinline|#1|}  % inline source code
\newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}
\newcommand{\DRbar}{\textoverline{DR}\xspace}
\newcommand{\MSbar}{\textoverline{MS}\xspace}
\newcommand{\unit}[1]{\,\text{#1}}      % units
\newcommand{\userinput}{\text{input}}
\newcommand{\pole}{\text{pole}}
\newcommand{\Lagr}{\mathcal{L}}
\newcommand{\unity}{\mathbf{1}}
\newcommand{\figref}[1]{\figurename~\ref{#1}}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\tabref}[1]{\tablename~\ref{#1}}
\newcommand{\ptitle}[1]{\emph{#1}}
\renewcommand{\ptitle}[1]{}
\newcommand{\Zv}{\mathbf{\backslash}\mkern-11.0mu{Z}}
\makeatletter
\lstnewenvironment{numlstlisting}[1][]{%
  \lstset{%
    #1,
    numbers=left,
    firstnumber=auto,
    numberstyle=\tiny\sffamily}%
  \csname\@lst @SetFirstNumber\endcsname
}{%
  \csname \@lst @SaveFirstNumber\endcsname
}
\makeatother
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\re}{Re}
\DeclareMathOperator{\im}{Im}
\def\at{\alpha_t}
\def\ab{\alpha_b}
\def\as{\alpha_s}
\def\atau{\alpha_{\tau}}
\def\oat{O(\at)}
\def\oab{O(\ab)}
\def\oatau{O(\atau)}
\def\oatab{O(\at\ab)}
\def\oatas{O(\at\as)}
\def\oabas{O(\ab\as)}
\def\oatababq{O(\at\ab + \ab^2)}
\def\oatqatababq{O(\at^2 + \at\ab + \ab^2)}
\def\oatasatq{O(\at\as + \at^2)}
\def\oatasabas{O(\at\as +\ab\as)}
\def\oatasabasatq{O(\at\as + \at^2 +\ab\as)}
\def\oatq{O(\at^2)}
\def\oabq{O(\ab^2)}
\def\oatauq{O(\atau^2)}
\def\oabatau{O(\ab \atau)}
\def\oatplusabsq{O((\at+\ab)^2)}
\def\oas{O(\as)}
\def\oatauqatab{O(\atau^2 +\ab \atau )}

@@

\subsection{Current limitations and future extensions}
Although we try to handle as many models as possible there are still
some limitations to what can be done. In it's current form \fs assumes
all couplings are are real, therefore it is limited to $CP$-conserving
versions of SUSY models.   Although it is implicit in the
title we would like to stress that currently we cannot provide a
spectrum generator for non-SUSY models.  This also means that like
other spectrum generators it is difficult to reliably predict the mass
spectrum in extreme cases such as Split-SUSY
\cite{Wells:2003tf,ArkaniHamed:2004fb,Giudice:2004tc} where the mass
scales of the SUSY model are drastically split leading to very large
logarithms which are not resummed.  In such cases the tower of
effective theories offers the best possibility for a solution.
However for this to work first \fs must be extended to include
non-SUSY models.  Finally the gauge group of the model should be
semi-simple and contain the standard model gauge group as factor so
that we can match to the SM at low energies.

@@

\fs allows the user to add leading two-loop contributions to the
$CP$-even and $CP$-odd Higgs self-energies as well as to the $CP$-even Higgs
tadpoles.  For MSSM-like models (with two $CP$-even Higgs bosons, one
$CP$-odd Higgs boson, one neutral Goldstone boson) routines dor
caluclating these corrections will be generated by setting
\code{UseHiggs2LoopMSSM = True} in the model file and by defining the
effective $\mu$-term \code{EffectiveMu = \\[Mu]}.  This will add the
zero-momentum corrections of the order $O(y_t^4 + y_b^2 y_t^2 +
y_b^4)$, $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$, $O(y_\tau^4)$ from
\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}.
For NMSSM-like models (with three $CP$-even Higgs bosons, two $CP$-odd
Higgs bosons, one neutral Goldstone boson) the two-loop contributions
are generated by setting \code{UseHiggs2LoopNMSSM = True} and by
defining the effective $\mu$-term like \code{EffectiveMu = \\[Lambda]
  vS / Sqrt[2]}, for example.  This will add the zero-momentum
corrections of the order $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ from
\cite{Degrassi:2009yq} and also the MSSM contributions of the order
$O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$ as well
\cite{Brignole:2001jy,Dedes:2003km} which only represent a partial
correction for that order in the NMSSM, but can be a good
approximation when singlet mixing is very small\footnote{These
  corrections may be disabled in the SLHA file, as described in
  section \secref{sec:PoleMasses}.}.

@@

The CKM mixing matrix is currently set to unity and $CP$-violating

@@

(i) the calculated lightest $CP$-even Higgs pole mass, $m_h$, differs

@@

Since the Higgs mass is a very important measurement and the two-loop
corrections can be larger than the current experimental error
\cite{Degrassi:2009yq} we recommend to set these switches to
\code{True} in any MSSM-\footnote{In \fs a model is MSSM-like if (i)
  its superpotential is approximately given by the MSSM one, (ii) it
  implements $R$-parity conservation or something equivalent, and
  (iii) contains two $CP$-even and $CP$-odd Higgs bosons, where one $CP$-odd
  Higgs boson may be a Goldstone boson, all with an MSSM-like coupling
  to $t$, $b$ and $\tau$.} and NMSSM-like\footnote{In \fs a model is
  NMSSM-like if (i) its superpotential is approximately given by the
  NMSSM one, (ii) it implements $R$-parity conservation or something
  equivalent, and (iii) contains three $CP$-even and $CP$-odd Higgs
  bosons, where one or two $CP$-odd Higgs bosons may be Goldstone
  bosons, all with an NMSSM-like coupling to $t$, $b$ and $\tau$\@.
  Examples for NMSSM-like models are the USSM and the \ESSM.} model.
However in such models the user should still consider whether these
corrections are really the leading corrections in the model or if
there are other potentially large two-loop corrections which are
missing.  For models with a more extended Higgs sector we recommend
that the leading log two-loop corrections are estimated by
generalizing those of the MSSM or NMSSM.

@@

To check the correctness of \fs's generated spectrum generators
extensive unit testing against Softsusy's MSSM and NMSSM
implementations (both $Z_3$-invariant and $Z_3$-violating variants)
has been carried out.  These unit tests systematically compare all
tree-level mass matrices, EWSB equations, one- and two-loop
$\beta$-functions, one- and two-loop self-energies and one- and
two-loop tadpoles numerically for the CMSSM and NMSSM parameter points
given in \tabref{tab:unit-test-parameter-points}.
%
\begin{table}[tbh]
  \centering
  \begin{tabularx}{\textwidth}{lX}
    \toprule
    SUSY Model & Tested parameter points\\
    \midrule CMSSM & $m_0 = 125\unit{GeV}$, $M_{1/2} = 500\unit{GeV}$,
    $\tan\beta = 10$, $A_0 = 0$, $\sign\mu = \pm 1$
    \\
    $Z_3$-NMSSM & $m_0 = \{250,300\}\unit{GeV}$, $M_{1/2} =
    200\unit{GeV}$, $\tan\beta = 10$, $A_0 = -500\unit{GeV}$,
    $\sign\mu_\text{eff} = +1$, $\lambda = 0.1$
    \\
    $\Zv_3$-NMSSM & $m_0 = 540\unit{GeV}$, $M_{1/2} = 200\unit{GeV}$,
    $\tan\beta = 10$, $A_0 = -350\unit{GeV}$, $\sign\mu_\text{eff} =
    \pm 1$, $\lambda = \kappa = 0.1$, $s = 1\unit{TeV}$, $\mu' =
    290\unit{GeV}$, $m_S^{\prime 2} = 400\unit{GeV}$, $\xi_F =
    300\unit{GeV}$
    \\
    \bottomrule
  \end{tabularx}
  \caption{CMSSM and NMSSM parameter points used for the unit tests
    against Softsusy.
    We follow the notation of the NMSSM model parameters used in
    \cite{Ellwanger:2009dp,Allanach:2013kza}.}
  \label{tab:unit-test-parameter-points}
\end{table}
%
All tested expressions were found to agree within double machine
precision.\footnote{Due to the systematic and detailed tests several
  bugs in Softsusy, SARAH and \fs could be identified and corrected.}
Furthermore, the output of the iterative procedures which solve the
one- and two-loop corrected tadpole equations
\eqref{eq:one-loop-ewsb-eq} to find the minimum of the effective Higgs
potential were compared numerically for these parameter points and
found to agree within machine precision as well.  Finally, the overall
pole mass spectrum and mixing after the full fixed-point iteration has
finished has been compared, and was found to agree at the sub-permille
level.  The origin of the sub-permille level difference between \fs
and Softsusy is the different determination of the weak mixing angle
$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ in the SUSY model in the
\DRbar\ scheme: \fs calculates
$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from $M_W$ and $M_Z$, as
described in \secref{sec:calculation-of-gauge-couplings}, while
Softsusy determines $\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from
the muon decay constant $G_\mu$.  The approach used in Softsusy
results in more precise \DRbar\ gauge couplings at the $M_Z$ scale,
because the muon decay constant is known with a higher accuracy than
the $W$-boson mass.  Furthermore, in Softsusy $3.5.0$ some three-loop
$\beta$-functions and two-loop threshold corrections can be enabled in
the MSSM to increase the accuracy of the RG running and the
determination of the \DRbar\ gauge and Yukawa couplings at $M_Z$
\cite{Allanach:2014nba}.  These corrections are not implemented in \fs
so far.  The complete set of unit tests is shipped with \fs and can be
found in the \code{test/} directory.  The tests can be run with the
command \code{make execute-tests}.  All unit tests are carried out
nightly in order to continuously check the correctness of the meta
code and the generated spectrum generators for the shipped models.
The nightly test results can be found at
\url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/test.xhtml}

One of \fs's design goals is a short run-time.  In this section we
demonstrate that this goal was achieved by comparing the run-time of
two different sets of CMSSM spectrum generators:
%
\begin{itemize}
\item \emph{Without sfermion flavour violation:} Disallowing sfermion
  flavour violation simplifies the calculation of the pole masses,
  because flavour-off-diagonal sfermion self-energy matrix elements
  don't need to be calculated.  Here we compare \fs's non-flavour
  violating CMSSM spectrum generator FlexibleSUSY-NoFV (version 1.0.0)
  against SPheno (version 3.2.4) and Softsusy (version 3.4.0).
%
\item \emph{With sfermion flavour violation:} Allowing for sfermion
  flavour violation in general increases the run-time of spectrum
  generators, because the full $6\times 6$ sfermion self-energy
  matrices have to be calculated.  Here we compare FlexibleSUSY-FV
  (version 1.0.0) and SPhenoMSSM (generated with SARAH 4.1.0 and
  linked against SPheno 3.2.4).  Both spectrum generators are based on
  SARAH's MSSM model file, which allows for sfermion flavour
  violation.
\end{itemize}
%
For the run-time comparison the CKM matrix is set to unity, all
$CP$-violating phases are set to zero and $R$-parity violation is
disabled.  \fs and Softsusy are compiled with g++ 4.8.0 and Intel
ifort 13.1.3 20130607.  SPheno and SPhenoMSSM are compiled with Intel
ifort 13.1.3 20130607.\footnote{Intel's ifort compiler decreases the
  run-time of SPheno and SPhenoMSSM by approximately a factor $1.5$,
  compared to gfortran.}  We're generating $2\cdot 10^{4}$ random
CMSSM parameter points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
[50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
and $A_0\in [-1000,$ $1000]\unit{GeV}$.  For each point an SLHA input
file is created by appending the values of $m_0$, $m_{1/2}$,
$\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
SLHA template file given in \ref{sec:speed-test-slha-template-file}.
The resulting SLHA input file is passed to each spectrum generator and
the (wall-clock) time is measured until the program has finished.  The
average run-times for three different CPU types can be found in
\tabref{tab:run-time-comparison}.  The first column shows the run-time
on a Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
enabled.  The second column shows the run-time on the same processor
where both cores were enabled.  In the third column a machine with two
Intel Xeon CPUs (L5640, $2.27\unit{GHz}$, $6$ cores) was used.
%
\begin{table}[tbh]
  \centering
  \begin{tabular}{llll}
    \toprule
                            & Intel Core2 Duo    & Intel Core2 Duo   & $2$ $\times$ Intel Xeon\\
                            & (P8600, $1$ core)  & (P8600, $2$ cores)& (L5640, $6$ cores)\\
    \midrule
    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
    \midrule
    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
    SPhenoMSSM 4.1.0        & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
    \bottomrule
  \end{tabular}
  \caption{Average run-time of CMSSM spectrum generators
    for random parameter points.  The first three rows show three
    spectrum generators with disabled sfermion flavour violation.  Rows
    $4$--$5$ show two spectrum generators with enabled sfermion flavour violation, both based
    on SARAH's MSSM model file.}
  \label{tab:run-time-comparison}
\end{table}

\section{Examples of C++ code adaptation}
\label{sec:examples of c++ code adaptation}

\subsection{Stacking models in a tower of effective theories}
\label{sec:tower construction}

@@

\subsection{Integrating custom-built C++ components}
\label{sec:integrating-custom-built}

@@
